On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-29 4_Result.tex
additional review (conclusions)

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Comma Operator                  & 0  & 100    & 0     \\
         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
         Post-Increment                  & 20 & 73.33  & 6.67  \\
         Omitted Curly Braces            & 0  & 100    & 0     \\
         Assignment as Value             & 20 & 60     & 20    \\
         Implicit Predicate              & 20 & 73.33  & 6.67  \\
         Logic as Control Flow           & 20 & 60     & 20    \\
         Ternary Operator                & 60 & 26.67  & 13.3  \\
         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
         \textsc{overall}                & 18 & 71.33  & 8.64  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern.

@@

Six practitioners regarded the JavaScript language as easy to learn, while four
participants mentioned that the abundance of frameworks,
coupled with the abstractions they provide, make it so that one
can develop quickly in the language. However, as mentioned by one
of the interviewees, ``\emph{Developers who have started
programming in JavaScript might miss out on important lower-level
aspects, such as garbage collection, memory management,
and process management.}''.

Moreover, such abstractions might lead a developer to start to program
using the language without actually understanding some of its basic aspects.
For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
%% a lack of understanding of the languages core
%% features may be a problem worth investigating in future studies.

>>>>

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \emph{with} (LHS) and
      \emph{without} atoms of confusion (RHS).
      Participants were also allowed to choose NEUTRAL when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{LHS}}
                                      &  \multicolumn{1}{c}{\textsc{RHS}}
                                               & \multicolumn{1}{c}{NEUTRAL} \\ \midrule
         Comma Operator                  & 0  & 100    & 0     \\
         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
         Post-Increment                  & 20 & 73.33  & 6.67  \\
         Omitted Curly Braces            & 0  & 100    & 0     \\
         Assignment as Value             & 20 & 60     & 20    \\
         Implicit Predicate              & 20 & 73.33  & 6.67  \\
         Logic as Control Flow           & 20 & 60     & 20    \\
         Ternary Operator                & 60 & 26.67  & 13.3  \\
         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \midrule
         \textsc{overall}                & 18 & 71.33  & 8.64  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

Two other developers implicitly touched upon the callback pattern, mentioning that it can be difficult to understand \emph{asynchronous} programming in JavaScript. Since what is really happening at a lower level is that the JavaScript engine creates a callback stack that is separate from the main execution stack, and that callback functions are only executed when the main execution stack is empty, the concepts of asynchronous events and callbacks are inseparable in the language, and any abstractions for callback functions, such as promises and async/await syntax only hide the pattern. We also found other idioms that are JavaScript atom candidates, including
\emph{property access} via array subscription (\lstinline[language=javascript]{V1['P1']} instead of \lstinline[language=javascript]{V1.P1}) and \emph{arrow functions} (see listings~\ref{arr1} and~\ref{arr2}).

\begin{figure}
\begin{small}
\begin{lstlisting}[language=JavaScript,caption=Example of arrow function,label=arr1]
let inc = (x) => x + 1;
\end{lstlisting}
\begin{lstlisting}[language=JavaScript,caption=Alternative version without arrow function,label=arr2]
let inc = function(x) {
  return x + 1
}
\end{lstlisting}
\end{small}
\end{figure}

@@

% TODO: I don't see that our results support these claims.
% Let us comment for now.

%% Six practitioners regarded the JavaScript language as easy to learn, while four
%% participants mentioned that the abundance of frameworks,
%% coupled with the abstractions they provide, make it so that one
%% can develop quickly in the language. However, as mentioned by one
%% of the interviewees, ``\emph{Developers who have started
%% programming in JavaScript might miss out on important lower-level
%% aspects, such as garbage collection, memory management,
%% and process management.}''.

%% Moreover, such abstractions might lead a developer to start to program
%% using the language without actually understanding some of its basic aspects.
%% For instance, four participants argued that the language is not typed, two  believed the language runs exclusively on browsers, and two other developers stated that the language is natively asynchronous, all of which are incorrect statements.
%% For the multi-purpose language that is one of the most, if not the most used one by practitioners,
%% a lack of understanding of the languages core
%% features may be a problem worth investigating in future studies.

### 2021-04-29 5_Threats.tex
additional review (conclusions)

\textbf{Conclusion validity.} In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient. For the ones where we did not find statistically significant results for correctness, in fact the sample sizes we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these atoms are required, due to the likelihood of type II errors.

@@

\textbf{External validity.} Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
\castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

>>>>

\textbf{Conclusion validity.} In the context of our survey study, we apply different statistical tests appropriate for the cases where data was categorical (correctness, Chi-square test of independence) and continuous (time, Mann-Whitney U test). Furthermore, besides reporting p-values, we also report effect size measures appropriate for each scenario (odds-ratio and Cliff's delta) and apply a p-value correction technique to avoid the multiple testing problem. Finally, it could be argued that the size of the samples is insufficient to make conclusions for some of the atoms, a common problem in Empirical Software Engineering. We estimate the sample size for each one of the atom candidates, considering that each one had a different number of samples (Table~\ref{tab:difference-correctness}). To that end, we use the $\phi$ measure of effect size for each atom (based on the Chi-square statistic), the standard $\alpha$ coefficient of 0.05, and set the expected statistical power to 0.8, as usually employed in the literature~\cite{Ellis:2010:EGE}. As a result of this analysis, we discover that for all but one of the atoms where we had a statistically significant result for correctness, Implicit Predicate, the sample size is sufficient.
% For the ones where we did not find statistically significant results for correctness, in fact the sample sizes
% we have analyzed are insufficient for such small effect sizes. This indicates that further studies on these
% atoms are required, due to the likelihood of type II errors.

@@

\textbf{External validity.} Although our results suggest that the selected idioms and code constructs may lead to confusion for small code snippets, it is not clear whether that result extrapolates to other scenarios. Even though it is likely that in larger code bases the confusion induced by these constructs may be even greater, the existence of additional context may mitigate this effect. Another possible threat to the generalizability of our results, in particular for the survey and interviews, lies in the fact that the analyzed atoms may rarely occur in real JavaScript code bases. To mitigate that threat, we have analyzed 72 popular open source JavaScript repositories and found out that most of them are common, occurring at least once per 1,000 lines of code. Only Arithmetic as Logic and Comma Operator occur less frequently than once per 10,000 lines of code. 
% \castor{A forma como os átomos foram detectados é uma ameaça que não discutimos.}

### 2021-04-29 6_Conclusion.tex
additional review (conclusions)

\rb{draft. apenas movi esse texto que estava em settings para ca. pode servir de inspiracao para responder algumas das questoes de pesquisa.}

{\color{blue}Based on these results, we provide evidence that, although the atoms of confusion considered in our study do not take a much higher amount of time to predict the output of the code, their impact on program comprehension is, in almost all of our cases, highly significant.}

>>>>

This paper has reported the results of a mixed-method research effort to better understand the impact of atoms of confusion in JavaScript code. First, we conducted a survey with 140 JavaScript developers. We asked them to predict the output of some code snippets (with and without atoms). We provide evidences that seven atom candidates have a significant impact on program comprehension activities.

Afterwards, we interviewed 15 professional developers. In general, they considered code without atoms of confusion easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. Last but not least, we conducted a mining software repositories study. In particular, we found four atoms of confusion highly used in JavaScript code, i.e., Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment.

Our efforts have important implications for practice. First, we alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment, and Omitted Curly Braces). Second, our results might help tool developers to create program transformation tools to remove atoms of confusion automaticallu. They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).

As mentioned, the results of the interviews pointed out additional JavaScript constructs and idioms
that might introduce misunderstanding. These include not only nested callbacks and prototype-based
inheritance (which might not have a counterpart version), but also idioms such
as \emph{property access} and \emph{arrow functions}. As future work, we intend to
reproduce our study to validate whether or not these \emph{atom candidates} affect
the understanding of JavaScript code.

### 2021-04-29 main.tex
Updates from Overleaf
additional review (conclusions)

\usepackage{enumerate}
\usepackage{caption}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{supertabular}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{array}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xurl}
% \usepackage{appendix}
\usepackage[framemethod=TikZ]{mdframed}
\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 leftmargin       = 2,
 rightmargin      = 2,
 splittopskip     = \topskip]{mh}

@@

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
\lstset{
  basicstyle=\footnotesize\tt,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  language=Java,                 % the language of the code
  keywordstyle=\bf,
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  tabsize=2,                       % sets default tabsize to 2 spaces
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  morecomment=[s][\color{javadocblue}]{/**}{*/},
  numbers=left,
  numberstyle=\tiny\color{black},
  stepnumber=1,
}    
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present in {\color{red}7} out of \na atom candidates. In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

@@

\bibliographystyle{IEEEtran}
\bibliography{reference}

>>>>

\usepackage{enumerate}
\usepackage{balance}
\usepackage{caption}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{supertabular}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{colortbl}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{tabularx} 
\usepackage{multicol}
\usepackage{listings}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{array}
\usepackage{listings}
\usepackage{color}
\usepackage{scalefnt}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xurl}
% \usepackage{appendix}
\usepackage[framemethod=TikZ]{mdframed}
\definecolor{oldlace}{rgb}{0.99, 0.96, 0.9}
\newmdenv [ %
 skipabove=\topsep,
 skipbelow=\topsep,
 leftmargin       = 2,
 rightmargin      = 2,
 splittopskip     = \topskip]{mh}

@@

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % javadoc
\lstset{
  basicstyle=\footnotesize\tt,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  language=Java,                 % the language of the code
  keywordstyle=\bf,
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  tabsize=2,                       % sets default tabsize to 2 spaces
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  morecomment=[s][\color{javadocblue}]{/**}{*/},
  numbers=left,
  numberstyle=\tiny\color{black},
  stepnumber=1,
  keywords={let},
}    
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

\begin{abstract}
  Evolving legacy code is a challenging task, particularly when the code has been poorly written, increasing effort and imposing a significant cognitive load on developers. Although previous research has investigated the impact of small code patterns (hereafter atoms of confusion) that contribute to misunderstanding the source code written in \clang, \cpplang, and Java, little is known about the influence of these atoms on understanding dynamic languages (such as JavaScript). In this work, we investigate the impact of atoms of confusion on understanding code written in JavaScript in terms of a mixed-method research effort: a survey, a set of interviews with practitioners, and a mining software repositories (MSR) study. From our survey and interviews, we confirm that atoms of confusion also lead to code that is hard to understand in JavaScript: developers correctly predict the outcome of at least 15\% more cases when an atom of confusion is not present inseven out of ten atom candidates. In addition, our MSR effort reveals that atoms of confusion frequently arise in 72 popular open-source JavaScript systems: the Comma Operator is present in 11\% of the projects in our dataset; while five other atom candidates appear in more than 60\% of the projects, including the atoms Assignment as Value and Omitted Curly Braces, present in 8253 and 53357 cases, respectively. Altogether, our findings might help practitioners (a) better understand the implications of atoms of confusion on understanding JavaScript code, (b) avoid writing code that is unnecessarily difficult to maintain, and (c) design program transformation tools that remove potential sources of misunderstanding in JavaScript code.

@@

\balance
\bibliographystyle{IEEEtran}
\bibliography{reference}
